---
title: "a2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

read data
```{r}
getwd()
```

```{r}
setwd("C:/Users/Cryst/Documents/")
```

```{r}
data <- fread("http://www.secrepo.com/maccdc2012/conn.log.gz")
names(data) <- c("ts", "uid", "id.orig_h", "id.orig_p", "id.resp_h", "id.resp_p", "proto", "service", "duration", "orig_bytes", "resp_bytes", "conn_state", "local_orig", "missed_bytes", "history", "orig_pkts", "orig_ip_bytes", "resp_pkts", "resp_ip_bytes", "tunnel_parents" )
print(data[1:4,])
```

```{r}
summary(data)
```

```{r}
df = data
df$ts <- df$ts-min(df$ts)
print(min(df$ts))
```
Order the frame by time.
```{r}
df <- df[order(df$ts),]
print(df[dim(df)[1],'ts'] == max(df$ts)) # assess the order
```
Calculate the number of connects in each time stamp, choose each number as a state:
```{r}
ts_list <- unique(df$ts)
ts_count <- table(df$ts)
```

```{r}
ts_list
```

```{r}
ts_count
```
ts_count shows the number of connections at each time stamp,i.e. at time 62.3800001144409, 3 connections happened
calculate the numbers of each states:
```{r}
count_stat <- table(as.numeric(ts_count))
count_stat
```
Thus, there are state 1 are only hit 604229 times, state 2 are only hit 211377 times, ... only one time stamp have 729 connections, which is state 729. 
```{r}
1000/length(ts_count)
```
Omit the states hit less than 1000 times, which counts 0.052%, mark them as state 0, which means we only keep the state that less that 100 connection occured in a time stamp. the state space would be {0,1,2,...,100}.
```{r}
test<- as.data.frame(count_stat)
test_less <- test[which(test$Freq<1000),]
test_less
```

```{r}
less_states <- test_less$Var1
```

```{r}
ts_chain <- as.data.frame(ts_count)
ts_chain
```

```{r}
for(i in 1:nrow(ts_chain)){
  if(ts_chain$Freq[i] %in% less_states){
    ts_chain$Freq[i]<- 0
  } 
}
```

```{r}
ts_chain
```

```{r}
install.packages("markovchain")
library(markovchain)
verifyMarkovProperty(ts_chain$Freq)
```
Since the p-value shown is 0, we reject the null hypothesis that the sequence follows the Markov property. 
```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
the code with problem is HERE
```{r}
N = list()
t = list()
i0 = 1
for(i in c(1:dim(df)[1])){
  n0 = df[i0, "ts"]
  if(df[i, "ts"] != n0){
    N <- append(N, i-i0)
    t <- append(t, n0)
    i0 <- i
  }
}

```

test the code in a smaller scale of data
```{r}
df1 = df[1:100,]
N = list()
t = list()
i0 = 1
for(i in c(1:dim(df1)[1])){
  n0 = df1[i0, "ts"]
  if(df1[i, "ts"] != n0){
    N <- append(N, i-i0)
    t <- append(t, n0)
    i0 <- i
  }
}
```

```{r}
print(N)
```

```{r}
dim(df1[1])
```

```{r}

```

```{r}

```