---
title: "a2-markov-based-on-snort-ip"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

read data
```{r}
setwd("C:/Users/Cryst/Documents/")
```

```{r}
library(data.table)
```

```{r}
data <- fread("http://www.secrepo.com/maccdc2012/conn.log.gz")
names(data) <- c("ts", "uid", "id.orig_h", "id.orig_p", "id.resp_h", "id.resp_p", "proto", "service", "duration", "orig_bytes", "resp_bytes", "conn_state", "local_orig", "missed_bytes", "history", "orig_pkts", "orig_ip_bytes", "resp_pkts", "resp_ip_bytes", "tunnel_parents" )
print(head(data))
```

```{r}
snort1 <- read.csv("out1.csv")
snort2 <- read.csv("out2.csv")
snort3 <- read.csv("out3.csv")

```

```{r}
snort1
```

```{r}
snort2
```

```{r}
snort3
```

```{r}
data_snort2 <- data[snort2$x]
```

```{r}
unique(data_snort2$id.orig_h)
```
All attacks are from 10 originating ip address. 
```{r}
unique(data_snort2[,c("id.orig_h", "id.resp_h")])
```
and has 58 (originating, responding) ip address paires.
```{r}
data_snort2[,c("id.orig_h", "id.resp_h")]
data2<-cbind(data_snort2,snort2$y)
```

```{r}
print("table of originating ips in data2:")
table(data2$id.orig_h)
print("table of Responding ips in data2:")
table(data2$id.resp_h)
```

```{r}
4855/4975
```
In snort2, 97.59% attacks are from the ip address "192.168.202.79", 
```{r}
print("table resopnding ip address attacked by \"192.168.202.79\": ")
table(data2[which(data2$id.orig_h == "192.168.202.79")]$id.resp_h)
```
take the responding ip address as a process, assess whether it is a Markov chain:
```{r}
library(markovchain)
verifyMarkovProperty(data2[which(data2$id.orig_h == "192.168.202.79")]$id.resp_h)
#verifyMarkovProperty(data2$id.resp_h)
```

```{r}
#verifyMarkovProperty(data2[which(data2$id.orig_h == "192.168.202.79")]$V2)
verifyMarkovProperty(data2$V2)
```
Thus, both originating ip address and attack type processes are not Markov Process
Now verifying them in snort3
```{r}
data_snort3 <- data[snort3$x]
```

```{r}
unique(data_snort3$id.orig_h)
```
All attacks are from 5 originating ip address. 
```{r}
unique(data_snort3[,c("id.orig_h", "id.resp_h")])
```

and has 15 (originating, responding) ip address paires.
```{r}
data_snort3[,c("id.orig_h", "id.resp_h")]
data3<-cbind(data_snort3,snort3$y)
```

```{r}
print("table of originating ips in data3:")
table(data3$id.orig_h)
print("table of Responding ips in data3:")
table(data3$id.resp_h)
```

```{r}
11529/nrow(data3)
```
In snort3, 99.9% attacks are from the ip address "192.168.202.79", 
```{r}
print("table resopnding ip address attacked by \"192.168.202.79\": ")
table(data3[which(data3$id.orig_h == "192.168.202.79")]$id.resp_h)
```
take the responding ip address as a process, assess whether it is a Markov chain:
```{r}
library(markovchain)
verifyMarkovProperty(data3[which(data3$id.orig_h == "192.168.202.79")]$id.resp_h)
#verifyMarkovProperty(data3$id.resp_h)
```

```{r}
verifyMarkovProperty(data3[which(data3$id.orig_h == "192.168.202.79")]$V2)
#verifyMarkovProperty(data3$V2)
```
Thus, both originating ip address and attack type processes are not Markov Process.
```{r}
data_snort <- data[c(snort1$x, snort2$x, snort3$x)]
```

```{r}
unique(data_snort$id.orig_h)
```

```{r}
unique(data_snort[,c("id.orig_h", "id.resp_h")])
```

```{r}
snort <- rbind(snort1, snort2, snort3)
data0<-cbind(data_snort,snort$y)
data0
```

```{r}
print("table of originating ips in snort:")
table(data0$id.orig_h)
print("table of Responding ips in snort:")
table(data0$id.resp_h)
```

```{r}
16387/nrow(data0)
```

In snort, 99.2% attacks are from the ip address "192.168.202.79", 
```{r}
print("table resopnding ip address attacked by \"192.168.202.79\": ")
table(data0[which(data0$id.orig_h == "192.168.202.79")]$id.resp_h)
```
take the responding ip address as a process, assess whether it is a Markov chain:
```{r}
library(markovchain)
verifyMarkovProperty(data0[which(data0$id.orig_h == "192.168.202.79")]$id.resp_h)
verifyMarkovProperty(data0$id.resp_h)
```
Thus, both originating ip address and attack type processes are not Markov Process.

```{r}
full_data <- read.csv("ProcessedData.csv")
print(head(full_data))
library(markovchain)
verifyMarkovProperty(full_data$conn_type)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
