---
title: "04-Markov-Property"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
in this report, we will verify the Markov property of: "conn_type", "orig_ip", "orig_port", number of connections on each time stamp (denote as $N_t$), number of malicious connections on each time stamp (i.e. $N_t$ | conn_type == 1)


read data
```{r}
setwd("C:/Users/Cryst/Documents/")
```

```{r}
library(data.table)
```

```{r}
full_data <- read.csv("ProcessedData.csv")
```

```{r}
head(full_data)
```

```{r}
library(markovchain)
```
## 1.try to fit "conn_type" into markov model
```{r}
unique(full_data$conn_type)
```

```{r}
verifyMarkovProperty(full_data$conn_type)
```

## 2.try to fit ip addresses into markov model:

```{r}
sum(is.na(full_data$orig_ip)==TRUE)
```
no missing data in "orig_ip" column
```{r}
length(unique(full_data$orig_ip))
```
there are 180 unique orginating ip address in total, now count the times of each address shows:
```{r}
head(table(full_data$orig_ip))
```


no orginating ip address is missing 
```{r}
1000/length(full_data$orig_ip)
```
find out the address that occur less than 1000 times(0.07% of tota connections), mark them as "other"
```{r}
orig_ip_table <- as.data.frame(table(full_data$orig_ip))
minor_ips <- orig_ip_table[which(orig_ip_table$Freq<1000),]
```

```{r}
minor_states <- minor_ips$Var1
orig_ip_chain <- as.data.frame(full_data$orig_ip)
names(orig_ip_chain)<-"orig_ips"
```

```{r}
for(i in 1:nrow(orig_ip_chain)){
  if(orig_ip_chain$orig_ips[i] %in% minor_states){
    orig_ip_chain$orig_ips[i]<- "other"
  } 
}
```

```{r}
verifyMarkovProperty(orig_ip_chain$orig_ips)
```
thus, originating ip address doesn't fit markov chain

now fit responding ip address in markov chain
```{r}
sum(is.na(full_data$resp_ip)==TRUE)
```
no missing data in "resp_ip" column
```{r}
length(unique(full_data$resp_ip))
```
there are 2715 unique orginating ip address in total, now count the times of each address shows:
```{r}
head(table(full_data$resp_ip))
```

```{r}
1000/length(full_data$resp_ip)
```
find out the address that occur less than 1000 times(0.07% of tota connections), mark them as "other"
```{r}
resp_ip_table <- as.data.frame(table(full_data$resp_ip))
minor_ips <- resp_ip_table[which(resp_ip_table$Freq<1000),]
```

```{r}
minor_states <- minor_ips$Var1
resp_ip_chain <- as.data.frame(full_data$resp_ip)
names(resp_ip_chain)<-"resp_ips"
```

```{r}
for(i in 1:nrow(resp_ip_chain)){
  if(resp_ip_chain$resp_ips[i] %in% minor_states){
    resp_ip_chain$resp_ips[i]<- "other"
  } 
}
```

```{r}
verifyMarkovProperty(resp_ip_chain$resp_ip)
```
thus, originating ip address doesn't fit markov chain

## 3.try to fit ports into markov model:

```{r}
sum(is.na(full_data$orig_port)==TRUE)
```
no missing data in "orig_port" column
```{r}
length(unique(full_data$orig_port))
```
there are 34309 unique orginating ip address in total, now count the times of each address shows:
```{r}
head(table(full_data$orig_port))
```

```{r}
1000/length(full_data$orig_ip)
```
find out the ports that occur less than 1000 times(0.07% of tota connections), mark them as "other"
```{r}
orig_port_table <- as.data.frame(table(full_data$orig_port))
minor_ports <- orig_port_table[which(orig_port_table$Freq<1000),]
```

```{r}
minor_states <- minor_ports$Var1
orig_port_chain <- as.data.frame(full_data$orig_port)
names(orig_port_chain)<-"orig_ports"
```

```{r}
for(i in 1:nrow(orig_port_chain)){
  if(orig_port_chain$orig_ports[i] %in% minor_states){
    orig_port_chain$orig_ports[i]<- "other"
  } 
}
```

```{r}
verifyMarkovProperty(orig_ip_chain$orig_ips)
```

## 4. try to fit the number of connections in each time stamp into markov chain
```{r}
df = full_data
min_ts = min(df$ts)
df$new_ts <- df$new_ts-min_ts
print(min(df$new_ts))
```

```{r}
df <- df[order(df$ts),]
print(df[dim(df)[1],'ts'] == max(df$ts)) # assess the order
```

Calculate the number of connects in each time stamp, choose the numbers as states:
```{r}
ts_list <- unique(df$ts)
ts_count <- table(df$ts)
```

```{r}
head(ts_count)
```

ts_count shows the number of connections at each time stamp,i.e. at time 0, there were 4 connections; at time 1.34999990463257, there was only 1 connection

calculate the numbers of each states:
```{r}
count_stat <- table(as.numeric(ts_count))
head(count_stat)
```
Thus, state 1 are hit 604229 times, state 2 are hit 211377 times, ... only one time stamp have 729 connections.
```{r}
1000/length(ts_count)
```
Omit the states hit less than 1000 times, which counts 0.052%, mark them as state 0, which means we only keep the state that less that 100 connections occurred in a time stamp. the state space would be {0,1,2,...,100}.

```{r}
temp<- as.data.frame(count_stat)
minor <- temp[which(temp$Freq<1000),]
minor
```
```{r}
minor_states <- minor$Var1
```

```{r}
ts_chain <- as.data.frame(ts_count)
ts_chain
```

```{r}
for(i in 1:nrow(ts_chain)){
  if(ts_chain$Freq[i] %in% minor_states){
    ts_chain$Freq[i]<- 0
  } 
}
```

```{r}
names(ts_chain) <- c("ts", "N_t")
ts_chain
```

```{r}
verifyMarkovProperty(ts_chain$Freq)
```
Since the p-value shown is 0, we reject the null hypothesis that the sequence follows the Markov property. 

5. try to fit the number of malicious connections in each time stamp into markov chain
```{r}
df = full_data[which(full_data$conn_type == 1),]
df$ts <- df$ts-min(df$ts)
print(min(df$ts))
```
Order the frame by time.
```{r}
df <- df[order(df$ts),]
print(df[dim(df)[1],'ts'] == max(df$ts)) # assess the order
```
Calculate the number of connects in each time stamp, choose each number as a state:
```{r}
ts_list <- unique(df$ts)
ts_count <- table(df$ts)
```

```{r}
head(ts_count)
```
ts_count shows the number of connections at each time stamp,i.e. at time 62.3800001144409, 3 connections happened
calculate the numbers of each states:
```{r}
count_stat <- table(as.numeric(ts_count))
count_stat
```
```{r}
ts_chain<-as.data.frame(ts_count)
verifyMarkovProperty(ts_chain$Freq)
```
Since the p-value shown is 0, we reject the null hypothesis that the sequence follows the Markov property. 


